Interacting Classical Gas

### 2016-10-02 Interacting Classical Gas_Zhiyu_APS.tex
Oct2 APS's PRE format, tex

\documentclass[a4paper,onecolumn,12pt]{article}
\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
 \usepackage{enumerate}
 \usepackage{verbatim}% for \begin{comment} \end{comment}
 \usepackage{color}
 \begin{document}
 \title{Interacting Classical Gas}
 \author{Zhiyu}
 \maketitle
 \tableofcontents
 \newpage

\section{Preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}

\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
It can be easily shown that eq.\ref{eq:preparation2} can be rewrite into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
with the transformation:
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\end{equation}

In this manner, we reduced number of the parameters in our model into three: $\tilde{H}$, $\tilde{F_0}$, $N$. Equivalently, we set $\sigma$, $\omega$ and $m$ to 1 in our numerical simulation. Throughout, we will use ``E" and ``$F_0$" to denote their reduced version  $\tilde{H}$, $\tilde{F_0}$.

\subsection{overview of the system}
\paragraph{relation between Radius and Energy}
radius R is defined as the mean-square-root of each particles position $\left\lbrace x_i\right\rbrace$. When E is very large, the interaction energy can be omitted, so E is dominated by the kinetic energy and potential energy of harmonic oscillators. Thus, $E=2*\frac{1}{2}\omega_0^2*\sum_{i}x_i^2=N\omega_0^2R^2$ 
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.5]{D:/OriginUserfiles/MPI_internship/E_R_fitting.jpg}
-\caption{E-R} This picture shows $E=kR^2$ when energy is large. ($k=N\omega_0^2$)
\end{figure}

\paragraph{Ground states}
For the ground state of this system, one could predict that \\(1) When the sigma is larger than the size of the cloud (in view of the scaling behaviour, large $\sigma$ is equivalent to \textbf{small F0 limit} when we fix $\sigma=1$), it can be easily shown that the ground state is crystal-like, since all every particle keep a equal distance with its nearest neighbours, which is $\frac{2F0}{\omega_0^2}$. The low-lying excitation is particles' independent harmonic oscillation around their equilibrium position.\\(2) When \textbf{F0 is very large} the molecules act like a hardcore gas, while the size of particle is just $\sigma$.

In the picture below, you can see this crossover from the ``solid-like" limit to the ``hardcore gas" limit.

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.3]{D:/matlab2016/MolecularDynamics/figure/groundstate1.eps}
\caption{particles' position at ground states }
\end{figure}

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.3]{D:/matlab2016/MolecularDynamics/figure/Eg1.eps}
\caption{ground state Energy with $F_0$}
\end{figure}

\newpage
\section{Breathing Frequency}
Usually, to see the response of a quantum gas to a quench, one start from the ground state and suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R oscillating at a certain frequency, which is usually called ``breathing mode".
In classical version, the ground state may be meaningless--since classical limit is $\bar{h}\rightarrow 0$, no matter how low the kT is, $kT/\bar{h}$ is always zero so that classical ground state is quite different from the quantum one. Instead, one will expect that the thermalized high temperature classical gas may have something in common with the high excited state of quantum gas. In view of this, if we want to do a quench similar to quantum case, we should begin with a thermalized distribution at $\omega_0'$ and evolve to a new thermalized state at varied $\omega_0$. Furthermore, notice that X and P (normalized) are symmetric for harmonic oscillator. If there is no spontaneous symmetry breaking in phase-space, the distribution cloud should be circular when the gas is completely thermalized. The difference between thermalized cloud before and after quench is  circular under different normalization of phase-space. If we look at both of them in the phase-space normalized according to $\omega_0$, we will find the former one is an ellipse while the latter one is a circle.
If the ellipse is only slightly deviated from circle, the oscillating amplitude will be too small to be distinguished from noise. Therefore, we start from an extreme case -- line distribution (random x, zero p) and observe the oscillation of R(t).
\begin{figure}[hbtp]
\center
\includegraphics[scale=0.8]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanF_differentE_log.eps}
\caption{breathing mode frequency measured at different E and $F_0$}
\label{Breathingfrequency1}
\end{figure}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.4]{D:/matlab2016/MolecularDynamics/figure/latexpic/rotatingframe_example2.eps} 
\includegraphics[scale=0.4]{D:/matlab2016/MolecularDynamics/figure/latexpic/rotatingframe_example1.eps} 
\caption{precession motion in phase-space observed in rotating frame (only show one or two particles' trajectory for clarity)}
\label{Breathingfrequency2}
\end{figure}

@@

@@

\section{Thermalization}
\subsection{Dynamics Study}
It is natural to think that a many-body system with interaction will be thermalized in ``usual" case. To study the condition of thermalization is equivalent to find out the mechanism that prohibits the system from ergodic. In view of this, we begin the study with two-particle motion. 
\paragraph{two-particle}

@@

We can always think of them in the reference frame of their center of mass ``C". In frame C, one can easily show that when there is no interaction, what we will see is that two particles are bound together by a harmonic trap center at C(cf). So their relative motion is also a harmonic oscillation.

@@

\paragraph{More particles}
Now let's consider the three-particle case. For two particle, the motion is non-chaotic. On the other hand, intuitively, we will say that three-body motion is chaotic so that the system could be ``thermalized" soon. However, in some cases, the time scale of thermalizing could still be very long. Suppose two of them, say, A and B, has small internal energy, which means their distance and relative velocity are both small. Meanwhile, suppose particle C has some energy quite different from A \& B. In this case, A \& B will often be in the interaction, while C will pass both of them at a high speed in each period. How will energy transfer between C and the two-particle system A and B? Since the relative velocity of C and the two-particle system is usually large, C will pass A-B pair in a short time $\tau\sim\frac{\sigma}{R}$ ($\tau\ll\frac{2\pi}{\omega}$). C gives A a push when approaching it, and then push A back when leaving.  As has been discussed in the two-particle case, this process is equivalent to giving A a very small speed($\sim O(\frac{\sigma}{R})$) in the background of a trap. Since both the position and velocity of A and B are close($|x_A-x_B|\sim\sigma$), the velocity increase of A \& B are almost the same (difference$\sim O(\frac{\sigma^2}{R^2})$). In this manner, the passing of particle C only kicks the center of mass of the A-B pair slightly, leaving the internal motion of the pair unaffected. In another word, the existence of C will not have significant effect on the energy transfer between A and B, but only ``dance" with their center of mass. The physics of the ``dance" is similar to the dance between two particle. As is shown in fig.\ref{fig:thermalization4}, the particles with energy close to each other tends to form a pair with small internal energy and the pair's internal energy transfer is relatively stable. %We only need to substitute $m$, $F_0$ \& $\sigma$ with their effective counterparts, e.g. change $m$ to $\mu$ , which is the reduced mass of m \& 2m.

@@

\paragraph{energy distribution}
For an isothermal system, the energy distribution of the whole system follows the Boltzmann distribution in equilibrium state. However, to our knowledge, there is not any obvious conclusion about the distribution for an isolated system where the total energy is conserved.

@@

\paragraph{Shape of Distribution in Phase-space}
If the oscillating phases of particles are highly correlated, for example, initial state with zero velocity, the shape of cloud will be a thin ellipse. If the phase of oscillation is completely random, the cloud should be isotropic in phase space. In view of this, we defined a shape factor S:
\\$S=\frac{a-b}{a+b}$, where a and b are long axis and short axis of inertia ellipse in phase space respectively. a,b can be calculated by diagonalizing the inertia matrices I: $I_{xx}=\sum{x^2}, I_{xp}=I_{px}=\sum{xp},I_{pp}=\sum{p^2}$
S=1 for line-shape distribution, S=0 for circular distribution.
The time evolution of S will probably show how fast the distribution ``forget" its initial shape.

>>>>

\documentclass[aps,pre,twocolumn
,groupedaddress]{revtex4-1}
%\documentclass[a4paper,onecolumn,12pt]{article}
\usepackage{amsmath}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{graphicx}
 \usepackage{subfigure}
 \usepackage{enumerate}
 \usepackage{verbatim}% for \begin{comment} \end{comment}
 \usepackage{color}
 \usepackage{bm}
 \begin{document}
 \title{Interacting Classical Gas}
 \author{Zhiyu}
 \affiliation{Fudan University}
\begin{abstract}
\end{abstract}
 \maketitle
% \tableofcontents
% \newpage
\section{Introduction}

\section{Preparation}
\subsection{Scaling behaviour}
We investigated the classical gas in harmonic trap with a repulsive interaction. We chose one of the simplest types of interaction:
\begin{equation}
F=\left\lbrace
\begin{aligned}
&F=F0 	&,x\textless\sigma\\
&F=0 	&,x\textgreater\sigma
\end{aligned}
\right.
\label{eq:preparation1}
\end{equation}
so the Hamiltonian is
\begin{equation}
H=\frac{1}{2}m\omega_0^2\sum_{i}x_i^2+\frac{1}{2}m\sum_i v_i^2+\sum_{\left|x_i-x_j\right|\textless\sigma}F_0\left(\sigma-\left|x_i-x_j\right|\right)
\label{eq:preparation2}
\end{equation}
It can be shown that eq.\ref{eq:preparation2} can be rewrite into the following form:
\begin{equation}
\tilde{H}=\frac{1}{2}\sum_{i}\tilde{x_i}^2+\frac{1}{2}\sum_i\left(\frac{d\tilde{x}}{d\tilde{t}}\right)^2+\sum_{\left|\tilde{x_i}-\tilde{x_j}\right|\textless 1}F_0\left(1-\left|\tilde{x_i}-\tilde{x_j}\right|\right)
\end{equation}
with the transformation:
\begin{equation}
\left\lbrace
\begin{aligned}
&\tilde{H}&=&\frac{H}{m\omega_0^2\sigma^2}\\
&\tilde{x_i}&=&\frac{x_i}{\sigma}\\
&\tilde{t}&=&\omega_0t\\
&\tilde{F_0}&=&\frac{F_0}{m\omega_0^2\sigma}\\
\end{aligned}
\right.
\end{equation}

In this manner, we reduced number of the parameters in our model into three: $\tilde{H}$, $\tilde{F_0}$, $N$. Equivalently, we set $\sigma$, $\omega_0$ and $m$ to 1 in our numerical simulation. Throughout, we will use ``E" and ``$F_0$" to denote their reduced version  $\tilde{H}$, $\tilde{F_0}$.

\subsection{Overview}
\subsubsection{Relation between Radius and Energy}
The radius of the cloud R is defined as the root-mean-square of particles' position $\left\lbrace x_i\right\rbrace$. When E is very large, the interaction energy can be omitted, so E is dominated by the kinetic energy and potential energy of harmonic oscillators. Thus, $E=\omega_0^2\sum_{i}x_i^2=N\omega_0^2R^2$ 
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.3]{D:/OriginUserfiles/MPI_internship/E_R_fitting.jpg}
-\caption{E-R} This picture shows $E=kR^2$ when energy is large. ($k=N\omega_0^2$)
\end{figure}

\subsubsection{Ground states}
For the ground state of this system, one could predict that \\(1) When the $\sigma$ is larger than the size of the cloud (in view of the scaling behaviour, large $\sigma$ is equivalent to \textbf{small $\mathbf{F_0}$ limit} when we fix $\sigma=1$), it can be easily shown that the ground state is crystal-like, since all every particle keeps an equal distance with its nearest neighbours, which is $\frac{2F_0}{\omega_0^2}$. The low-lying excitation is particles' independent harmonic oscillation around their equilibrium position.\\(2) When \textbf{$\mathbf{F_0}$ is very large} the molecules act like a hardcore gas, while the size of particle is just $\sigma$.

In the fig.\ref{fig:GS1}, we can see this crossover from the ``solid-like" limit (left) to the ``hardcore gas" limit (right).

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.5]{D:/matlab2016/MolecularDynamics/figure/latexpic/groundstate1.eps}
\caption{particles' position at ground states }
\label{fig:GS1}
\end{figure}

\begin{comment}
\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.3]{D:/matlab2016/MolecularDynamics/figure/Eg1.eps}
\caption{ground state Energy with $F_0$}
\end{figure}
\end{comment}

\newpage
\section{Breathing Frequency}
Usually, to see the response of quantum gas to a quench, one begin simulation with the ground state and then suddenly vary some parameter by a little, for instance, $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$. In that case, one will observe the radius R oscillating at a certain frequency, which is usually called ``breathing mode".
However, it is quite different in the classical version. Firstly, the ground state may be meaningless. In quantum case, the reason why we care about ground state and low lying excitations is that BEC occur at low temperature. But in classical case, there is no BEC. So it is not necessary to focus on low lying excitations in classical case. Besides, since classical limit is $\bar{h}\rightarrow 0$, no matter how low the kT is, $\bar{h}/kT$ is always zero so that classical behavious of low lying excitations is quite different from the quantum one. Secondly, it is no longer necessary to vary parameters ``by a little" in the quench. In quantum case, the small quench $\omega_0'\rightarrow\omega_0=\omega_0'+\Delta\omega$ means the spectrum is shifted slightly, thus one may expect to see some beat phenomenon({\color{red}{?}}). But in classical case, since $\bar{h}$ vanishes, no matter how small the quench is, one can never recover or proximate the quantum case. Thus it is not necessary to limit ourselves to small quench.

Furthermore, notice that X and P (normalized) are symmetric for simple harmonic oscillator. In weak interaction limit, the distribution cloud should be circular when the gas is completely thermalized. The difference between thermalized cloud before and after quench is that they are circular under different normalization of phase-space. If we look at both of them in the phase-space normalized according to the final $\omega_0$, we will find shape of the cloud before the quench is an ellipse while the final one is a circle.

Since the number of particles we use in the Molecular Dynamics simulation is limited (usually 5 to 20), the random noise is significant. If we start with a state whose configuration in phase-space is an ellipse slightly deviated from circle, the oscillating amplitude will be too small to be distinguished from noise. Therefore, we start from an extreme case -- line distribution (random x, zero p) and observe the oscillation of R(t). 
\begin{figure}[hbtp]
\center
\includegraphics[scale=0.6]{D:/matlab2016/MolecularDynamics/figure/latexpic/freq_scanF_differentE_log_2.eps}
\caption{breathing mode frequency measured at different E and $F_0$}
\label{Breathingfrequency1}
\end{figure}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.6]{D:/matlab2016/MolecularDynamics/figure/latexpic/rotatingframe_example2.eps} 
\includegraphics[scale=0.6]{D:/matlab2016/MolecularDynamics/figure/latexpic/rotatingframe_example1.eps} 
\caption{precession motion in phase-space observed in rotating frame (only show one(two) particles' trajectory in upper(lower) picture for clarity)}
\label{Breathingfrequency2}
\end{figure}

@@

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.25]{D:/matlab2016/MolecularDynamics/figure/latexpic/N=20_freq_at_differentF0_scanF0_1.eps}
\caption{•}
\end{figure}

@@

\section{Thermalization}
\subsection{Dynamics Study}
It is natural to think that a many-body system with interaction will be thermalized in ``usual" case. To study the condition of thermalization is equivalent to find out the mechanism that prohibits the system from ergodic. In view of this, we begin the study with two-particle motion. 
\subsubsection{two particles}

@@

We can always think of them in the reference frame of their center of mass ``C". In frame C, one can easily show that when there is no interaction what we will see is that two particles are bound together by a harmonic trap center at C(black line in fig.\ref{fig:thermalization2}). So their relative motion is also a harmonic oscillation.

@@

\subsubsection{More particles}
Now let's consider the three-particle case. For two particle, the motion is non-chaotic. On the other hand, intuitively, we will say that three-body motion is chaotic so that the system could be ``thermalized" soon. However, in some cases, the time scale of thermalizing could still be very long. Suppose two of them, say, A and B, has small internal energy, which means their distance and relative velocity are both small. Meanwhile, suppose particle C has some energy quite different from A \& B. In this case, A \& B will often be in the interaction, while C will pass both of them at a high speed in each period. How will energy transfer between C and the two-particle system A and B? Since the relative velocity of C and the two-particle system is usually large, C will pass A-B pair in a short time $\tau\sim\frac{\sigma}{R}$ ($\tau\ll\frac{2\pi}{\omega}$). C gives A a push when approaching it, and then push A back when leaving.  As has been discussed in the two-particle case, this process is equivalent to giving A a very small speed($\sim O(\frac{\sigma}{R})$) in the background of a trap. Since both the position and velocity of A and B are close($|x_A-x_B|\sim\sigma$), the velocity increase of A \& B are almost the same (difference$\sim O(\frac{\sigma^2}{R^2})$). In this manner, the passing of particle C only kicks the center of mass of the A-B pair slightly, leaving the internal motion of the pair unaffected. In another word, the existence of C will not have significant effect on the energy transfer between A and B, but only ``dance" with their center of mass. The physics of the ``dance" is similar to the dance between two particle. As is shown in fig.\ref{fig:thermalization4}, the particles with energy close to each other tends to form a pair with small internal energy and the pair's internal energy transfer is relatively stable. %We only need to substitute $m$, $F_0$ \& $\sigma$ with their effective counterparts, e.g. change $m$ to $\mu$ , which is the reduced mass of m \& 2m.

@@

\subsubsection{energy distribution}
For an isothermal system, the energy distribution of the whole system follows the Boltzmann distribution in equilibrium state. However, to our knowledge, there is not any obvious conclusion about the distribution for an isolated system where the total energy is conserved.

@@

\subsubsection{Shape of Distribution in Phase-space}
If the oscillating phases of particles are highly correlated, for example, initial state with zero velocity, the shape of cloud will be a thin ellipse. If the phase of oscillation is completely random, the cloud should be isotropic in phase space. In view of this, we defined a shape factor S:
\\$S=\frac{a-b}{a+b}$, where a and b are long axis and short axis of inertia ellipse in phase space respectively. a,b can be calculated by diagonalizing the inertia matrices I: $I_{xx}=\sum{x^2}, I_{xp}=I_{px}=\sum{xp},I_{pp}=\sum{p^2}$
S=1 for line-shape distribution, S=0 for circular distribution.
The time evolution of S will probably show how fast the distribution ``forget" its initial shape.
